---
title: "assignment_4"
author: "Alyssa Kibbe"
date: "11/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(lubridate)
library(kableExtra)
```

```{r}
hares <- read.csv(here("data", "bonanza_hares.csv"))
```


```{r}
juvenile <- hares %>%
  filter(age == "j") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% 
  mutate(day = day(date_new))

juvenile_per_year <- juvenile %>% 
  group_by(year) %>% 
  summarize(count_per_year = n())

ggplot(juvenile_per_year, aes(x =  year, y = count_per_year))+
  geom_col(fill = "goldenrod3",
          color = "goldenrod4" )+
  labs( x = "", y = "Hare Count", 
        title = "Variation in Juvenile Hare Count from 1999 to 2012")+
  theme_bw()


```


##### **Figure 1: Count of juvenile snowshoe hares over a 13 year period.** Populations counted at five sites along Bonanza Riparian zone on the Tanana River. 2002 and 2009 are excluded from the data set.  

### As seen in Figure 1, there is large variation in the snowshoe hare populations (min = 2 (2010), max = 126 (1999)). The overall trend of the data is skewed to the right. The mean number of snowshoe hare is 31 individuals counted per year. Median = 18.5. There are inconsistencies in the data that may have influences the results. In 2000, data was collected on 11 different days while in 2012, data was only collected on 2 days. For future collections, it would be beneficial to have a consistent number of sampling days each year.      

```{r}
# juvenile weights
ggplot(data = juvenile, aes(x = grid, y = weight))+
  geom_boxplot(aes(color = sex), )+
  labs(x = "Sample Site", y = "Weight (grams)", title = "Juvenile Snowshoe Hare Weight at Three Sampling Sites", colour = "Sex")
```

##### **Figure 2: Variation in weight between males and females at three sampling sites.** "bonbs" is  the sampling site consisting of lowland black spruce, "bonmat" or Bonanza Mature is a forested area filled with mature tree species. "bonrip" or Bonanza Riparian is the flood zone from the Tanana River. NA represents individuals with unidentified sex.

```{r}
ggplot(data = juvenile, aes(x = grid, y = weight))+
  geom_jitter(aes(color = grid), alpha = 0.7)+
  labs(x = "Sample Site", y = "Weight (grams)", title = "Juvenile Snowshoe Hare Weight Distribution by Sex", colour = "Sampling Site")+
  facet_wrap(~sex)
```



##### **Figure 3: Weight distribution of males and females at three sampling sites.** "bonbs" is  the sampling site consisting of lowland black spruce, "bonmat" or Bonanza Mature is a forested area filled with mature tree species. "bonrip" or Bonanza Riparian is the flood zone from the Tanana River.NA represents individuals with unidentified sex.

#### As seen in Figures 2, juvenile male hares have an greater average weight than females. In addition, the weights across the three locations show similar trends.However, there is the greatest difference in average weight in the Bonanza Mature sample population. The Bonanaza Riparian zone has the greatest number of sample points and thus has that largest variation in hare weights. 


```{r}
juvenile_summary <- juvenile %>% 
  group_by(sex) %>% 
  summarize(mean_weight = mean(weight,na.rm=T),
            sd_weight = sd(weight, na.rm=T),
            sample_size = n())
kbl(juvenile_summary) %>% 
  kable_styling()
```






